[![](./images/lindens.jpg)](https://en.wikipedia.org/wiki/Tilia) An Application of LIME to a Random Forest

### 2019-03-01 presentations/graphics group/An Application of LIME to a Random Forest.Rmd
some quick updates

1. Explanation of LIME 
2. Hamby bullet data
3. Applying LIME to the random forest
4. Issues with LIME and attempts at a solution
6. Future work
]

@@

[![](./images/blackbox.png)](https://en.wikipedia.org/wiki/Black_box)

@@

### Enter LIME

@@

- Default method in R package is gower distance
    - between 0 and 1 
    - 0 indicates exactly the same
- Other distance metrics can be specified

@@

- Since features are standardized, we can compare the coefficients
- Largest $\beta_i$ is conisdered the most important

@@

Then we tried more of the feature distribution estimaton methods
- explanations still not great
- $R^2$ values still very low
- predictions from the simple model still poor
- explanations dependent on the input

@@

- 2 to 5 quntile bins
- 2 to 5 equally spaced bins
- 2 to 5 tree based bins with `samesoure` as the response
- 2 to 5 tree based bins with `rfscore` as the response
- normal approximation
- kernel density estimation

@@

@@

#### Key Observations

- equally spaced bins are dependent on the nubmer of bins
- tree based bins usually have the lowest MSEs
- rfscore tree based bins usually have the highest $R^2$ for the binning options
- all options results in low $R^2$ values
    - we think that the linear regression is not similar enough to a random forest to produce good explanations
- explanations are relatively consistent across reps for first and second features
- non-binning methods produce very different results

@@

>>>>

1. Explanation of LIME 
2. Hamby bullet data
3. Applying LIME to the random forest
4. Issues with LIME and attempts at a solution
6. Conclusions and future work
]

@@

![](./images/blackbox.png)

@@

### Enter LIME...

@@

- Default method in R package is Gower distance
    - between 0 and 1 
    - 0 indicates exactly the same
- Other distance metrics can be specified

@@

- Since features are standardized, we can compare the coefficients
- Largest $\beta_i$ is considered the most important

@@

Then we tried more of the feature distribution estimation methods
- explanations still not great
- $R^2$ values still very low
- predictions from the simple model still poor
- explanations dependent on the input

@@

- 2 to 5 quantile bins
- 2 to 5 equally spaced bins
- 2 to 5 tree based bins with `samesoure` as the response
- 2 to 5 tree based bins with `rfscore` as the response
- normal approximation
- kernel density estimation

@@

class: middle, center, inverse

# Conclusions and Future Work

@@

- Equally spaced bins are dependent on the number of bins
- Tree based bins usually have the lowest MSEs
- rfscore tree based bins usually have the highest $R^2$ for the binning options
- All options results in low $R^2$ values
    - We think that the linear regression is not similar enough to a random forest to produce good explanations
- Explanations are relatively consistent across reps for first and second features
- Non-binning methods produce very different results

@@

- Apply LIME to explain a logistic regression model
    - We know how to interpret a logistic regression
    - Can compare to LIME explanations
- Run some sort of simulation to assess LIME explanations
- Adjust and compare weighting methods and the features selection methods
- Allow the number of bins to vary across features
- Use a tree as the simple model

# Image Credits

- slide 1: https://www.flickr.com/photos/lincolnian/300262799
- slide 2: https://ayoqq.org/explore/lime-drawing/
- slide 3: https://en.wikipedia.org/wiki/Black_box 
- slide 8: https://github.com/thomasp85/lime
